Speed and strength of an epidemic intervention

### 2020-02-28 doc/speedstrength.tex
Need to sleep

An epidemic can be characterized by speed (i.e., the exponential growth rate $r$) or strength (i.e., the reproductive number \RR).
Disease modelers have historically placed much more emphasis in strength, in part because the effectiveness of an intervention strategy is typically evaluated on this scale.  
Here we develop a mathematical framework for this classic, strength-based paradigm and show that there is a corresponding speed-based paradigm which can provide complementary insights.
In particular, we note that $r=0$ is a threshold for disease spread, just like $\RR=1$, and show that we can measure the speed and strength of an intervention on the same scale as the speed and strength of an epidemic, respectively.
We argue that, just as the strength-based paradigm provides the clearest insight into certain questions, the speed-based paradigm provides the clearest view in other cases.
\swp{We argue that the speed-based paradigm provides the clearest insight into certain questionsm just as the strength-based paradigm provides view in other cases.}
As an example, we show that evaluating the prospects of ``test-and-treat'' interventions \swp{(i.e., identifying infected individuals and putting them to clinical care, such as antiretroviral therapy)} against the human immunodeficiency virus (HIV) can be done more clearly on the strength than speed scale, given uncertainty in the proportion of HIV spread that happens early in the course of infection. \swp{you mean speed than strength, right?}
We suggest that disease modelers should refrain from putting too much emphasis on the reproductive number at the expense of the exponential growth rate, but instead look at these as complementary approaches.

@@

On the other hand, consider a ``test-and-treat'' strategy in which infected individuals are identified and put to clinical care, such as antiretroviral therapy (ART) \citep{garnett2009treating, granich2009universal, nah2017test}.
We assume that the rate of identification and removal is initially constant \swp{is approximately constant in the beginning}. 
Thus, the proportion continuing to transmit declines \swp{approximately} exponentially, and the effect of intervention $\tsub{L}{test}$ increases \swp{approximately} exponentially.
However, we assume that removal eventually saturates, to account for the effects of people who avoid identification, persistent treatment failures, and the possibility of rare transmission even under effective treatment
(\figref{test}A; see caption for the functional form of $\tsub{L}{test}$).
Under this assumption, the strength of intervention decreases with the proportion of early transmission in a manner roughly similar to the epidemic strength (\figref{test}B).
In this scenario, we predict that the intervention remains effective over the range of considered parameters.

>>>>

An epidemic can be characterized by speed (i.e., the exponential growth rate $r$) or strength (i.e., the reproductive number \RR).
Disease modelers have historically placed much more emphasis in strength, in part because the effectiveness of an intervention strategy is typically evaluated on this scale.  
Here we develop a mathematical framework for this classic, strength-based paradigm and show that there is a corresponding speed-based paradigm which can provide complementary insights.
In particular, we note that $r=0$ is a threshold for disease spread, just like $\RR=1$, and show that we can measure the speed and strength of an intervention on the same scale as the speed and strength of an epidemic, respectively.
We argue that, just as the strength-based paradigm provides the clearest insight into certain questions, the speed-based paradigm provides the clearest view in other cases.
\swp{We argue that the speed-based paradigm provides the clearest insight into certain questionsm just as the strength-based paradigm provides view in other cases.}
As an example, we show that evaluating the prospects of ``test-and-treat'' interventions against the human immunodeficiency virus (HIV) can be done more clearly on the speed than strength scale, given uncertainty in the proportion of HIV spread that happens early in the course of infection. 
We suggest that disease modelers should refrain from putting too much emphasis on the reproductive number at the expense of the exponential growth rate, but instead look at these as complementary approaches.

@@

On the other hand, consider a ``test-and-treat'' strategy in which infected individuals are identified and put to clinical care, such as antiretroviral therapy (ART) \citep{garnett2009treating, granich2009universal, nah2017test}.
We assume that the rate of identification and removal is approximately constant for newly infected individuals; thus, the proportion continuing to transmit declines approximately exponentially.
However, we assume that removal eventually saturates, to account for the effects of people who avoid identification, persistent treatment failures, and the possibility of rare transmission even under effective treatment
(\figref{test}A; see caption for the functional form of $\tsub{L}{test}$).
Under this assumption, the strength of intervention decreases with the proportion of early transmission in a manner roughly similar to the epidemic strength (\figref{test}B).
In this scenario, we predict that the intervention remains effective over the range of considered parameters.
